Current model cost effect

### 2020-09-23 src/current_model_effect.Rmd
error bootstrapping works

---
title: "Current model cost effect"
author: "Esa Turkulainen"
date: "8/25/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
#setwd("~/interval-sims/src")
library(tidyverse)
library(truncnorm)
library(caret)
library(boot)
source("helper_functions.R")
```

@@

The data object donation10 contains donation history data from beginning of 2010 to end of 2019 it has following columns.

@@

# Create subgroups from donation data
```{r subgroups}
# All successful whole blood donations
dons <- filter_donations(data = donation10, donat_phleb = "K", status = "-")
# Order donations by date
dons <- dons[with(dons, order(date)), ]
# Find only returning donors
rdons <- dons[duplicated(dons$donor),]

@@

# Find ratios
```{r}
# Male ratio
mr <- nrow(wmdon)/nrow(wdons)
# Female ratio
fr <- nrow(wfdon)/nrow(wdons)
# Male deferral ratio
mdr <- nrow(wmdef) / nrow(wdef)
# Female deferral ratio
fdr <- nrow(wfdef) / nrow(wdef)
# Total deferral prevalence
d <- nrow(wdef)/(nrow(wdef) + nrow(walldons))
```

The current model finds 1662/2116 = 78.5% of deferrals. In addition, it misclassifies 22.4% of donations as deferrals.
```{r}
# > conf
# Confusion Matrix and Statistics
# 
#           Reference
# Prediction Accepted Deferred
#   Accepted    31719      454
#   Deferred     9137     1662
TPR <- 1662/(1662 + 454) #78.5%
FPR <- 9137 / (31719 + 9137) # 22.4%
# Find q
q <- d * TPR
# Find a for 6 month category
# correctly predicted non-deferrals: 77.43% of the returning donations (all - d): (0.7743 * (1 - d))
# incorrectly predicted non-deferrals: 22.57% of the returning donations (all - d): (0.2257 * (1 - d))
a6 <- 1 * ((1 - FPR) * (1 - d)) + 
  3 * mr * (FPR * (1 - d)) + 2 * fr * (FPR * (1 - d)) + 
  3 * mdr * d + 2 * fdr * d

@@

# Final cost
```{r}
Pm <- 2 # estimate collected from FRCBS experts
Pd <- 60 # estimate collected from FRCBS experts
rl <- 0 # estimated previously from data
Fn = 0.1066 # etimated from data in https://github.com/FRCBS/interval-sims/blob/master/src/eba_sims_w_data.Rmd

@@

#What if we use some other than the default cut-off for model class assingment

@@

@@

```{r}
#count TPR and FPR for each probability
sen <- vector()
spe <- vector()
#df$obs has factor levels in wrong
#df$obs <- relevel(df$obs,"Deferred")

@@

get_cost <- function(TPR, FPR,d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) {
  q <- d * TPR
  # Find a for 6 month category
  # correctly predicted non-deferrals: 77.43% of the returning donations (all - d): (0.7743 * (1 - d))
  # incorrectly predicted non-deferrals: 22.57% of the returning donations (all - d): (0.2257 * (1 - d))
  a6 <- 1 * ((1 - FPR) * (1 - d.)) + 
    3 * mr. * (FPR * (1 - d.)) + 2 * fr. * (FPR * (1 - d.)) + 
    3 * mdr. * d. + 2 * fdr. * d.

@@

# > get_cost(TPR=0/2116, FPR=1- (0 /2116),d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) 
# # A tibble: 1 x 5
#       q    a6   a12    E6   E12
#   <dbl> <dbl> <dbl> <dbl> <dbl>
# 1     0  2.47  4.93  2.26  4.95
# > get_cost(TPR=2116/2116, FPR=1- (2116 /2116),d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) 
# # A tibble: 1 x 5
#        q    a6   a12    E6   E12
#    <dbl> <dbl> <dbl> <dbl> <dbl>
# 1 0.0327  1.04  1.11 -1.89 -1.76
#Yes, perfect model would give best savings

@@

@@

d <- tprs %>% ungroup() %>% select(probability,E6,E12) %>% pivot_longer(cols=c(E6,E12)) %>% mutate(name=gsub("^E",'',name))

size <- 2
p <- ggplot(d)
p <- p + geom_line(aes(x=probability,y=value,col=name)) + ylab("Cost effect (€ / donation)") + xlab("Probability of deferral cut-off") + labs(color="Deferral length (mo)")
p <- p + geom_point(aes( y =tprs$E6[which(tprs$E6 == min(tprs$E6))] , x = tprs$probability[which(tprs$E6 == min(tprs$E6))] ))
p <- p + geom_label(aes( y =tprs$E6[which(tprs$E6 == min(tprs$E6))] , x = tprs$probability[which(tprs$E6 == min(tprs$E6))] , label = paste0("p=",round(tprs$probability[which(tprs$E6 == min(tprs$E6))],2) , " ; ",  round(tprs$E6[which(tprs$E6 == min(tprs$E6))],2)," €")), nudge_y=0.2,size=size)
p <- p + geom_point(aes( y =tprs$E12[which(tprs$E12 == min(tprs$E12))] , x = tprs$probability[which(tprs$E12 == min(tprs$E12))] ))
p <- p + geom_label(aes( y =tprs$E12[which(tprs$E12 == min(tprs$E12))] , x = tprs$probability[which(tprs$E12 == min(tprs$E12))] , label = paste0("p=",round(tprs$probability[which(tprs$E12 == min(tprs$E12))],2) , " ; ",  round(tprs$E12[which(tprs$E12 == min(tprs$E12))],2)," €")), nudge_y=0.2,size=size)
p <- p + theme(legend.position="bottom")
p
ggsave(filename="../results/cme_prob_cut_curve.pdf", p, width = 90,  height = 80,units="mm", dpi=600, scale=1.0,encoding = "ISOLatin9.enc")
```

Using a bit heigher than the default cut-off  appears slightly better.
Let's bootstrap predictions to create error estimates for the predictions and subsequently on the cost effects.

# https://www.painblogr.org/2017-10-18-purrring-through-bootstraps.html
get_cost <- function(data, boot_ind){
  #classify predictions
  #get TPR and FPR
    conf <- caret::confusionMatrix(reference=df$obs, data=pred.classes[,i], positive = "Deferred",mode="everything")
    sen <- conf$byClass['Sensitivity'] 
    spe <- 1- conf$byClass['Specificity'] 
#gte cost
    get_cost(TPR=sen, FPR=spe,d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl)

#    return(coef(fit))
}

fit_boot <- boot(data, statistic= get_coefficients, R = nb_boot)
fit_boot_Bca <- compute_Bca_CI(fit_boot)

>>>>

---
title: "Current model cost effect"
author: "Esa Turkulainen & Mikko Arvas"
date: "`r Sys.time()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
#setwd("~/interval-sims/src")
library(tidyverse)
library(truncnorm)
library(caret)
library(boot)
source("helper_functions.R")
set.seed(123)
#echo "rmarkdown::render('current_model_effect.Rmd', clean=TRUE,output_format='pdf_document',output_file='../results/current_model_effect.pdf')" | R --slave
```

# Calculate parameters from donation history data

@@

The data object donation10 contains donation history data from beginning of 2010 to end of 2019. It has the following columns.

@@

## Create subgroups from donation data
```{r subgroups}
# All successful whole blood donations
dons <- filter_donations(data = donation10, donat_phleb = "K", status = "-")
# Order donations by date
dons <- dons[with(dons, order(date)), ]
# Find only returning donors
rdons <- dons[duplicated(dons$donor),]

@@

## Find ratios from data
```{r}
# Male ratio
mr <- nrow(wmdon)/nrow(wdons)
# Female ratio
fr <- nrow(wfdon)/nrow(wdons)
# Male deferral ratio
mdr <- nrow(wmdef) / nrow(wdef)
# Female deferral ratio
fdr <- nrow(wfdef) / nrow(wdef)
# Total deferral prevalence
d <- nrow(wdef)/(nrow(wdef) + nrow(walldons))
mr
fr
mdr
fdr
d
```

# Find interval adjustments from ratios and model metrics

The current model finds in validation data 1662/2116 = 78.5% of deferrals. In addition, it misclassifies 22.4% of donations as deferrals.
```{r}
# > conf
# Confusion Matrix and Statistics
# 
#           Reference
# Prediction Accepted Deferred
#   Accepted    31719      454
#   Deferred     9137     1662
TPR <- 1662/(1662 + 454) #78.5%
FPR <- 9137 / (31719 + 9137) # 22.4%
# Find q
q <- d * TPR
# Find a for 6 month category
# correctly predicted non-deferrals: 78.5% of the returning donations (all - d): (0.785 * (1 - d))
# incorrectly predicted non-deferrals: 22.4% of the returning donations (all - d): (0.224 * (1 - d))
a6 <- 1 * ((1 - FPR) * (1 - d)) + 
  3 * mr * (FPR * (1 - d)) + 2 * fr * (FPR * (1 - d)) + 
  3 * mdr * d + 2 * fdr * d

@@

# Final cost
```{r}
Pm <- 2 # estimate collected from FRCBS experts
Pd <- 60 # estimate collected from FRCBS experts
rl <- 0 # estimated previously from data
Fn = 0.1066 # estimated from donation history data in here https://github.com/FRCBS/interval-sims/blob/master/src/eba_sims_w_data.Rmd

@@

# What if we use some other than the default cut-off for model class assingment

@@

## Create different cuts

@@

# Count TPR and FPR for each probability cut-off

sen <- vector()
spe <- vector()

@@

get_cost <- function(TPR, FPR,d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) {
  q <- d * TPR
  a6 <- 1 * ((1 - FPR) * (1 - d.)) + 
    3 * mr. * (FPR * (1 - d.)) + 2 * fr. * (FPR * (1 - d.)) + 
    3 * mdr. * d. + 2 * fdr. * d.

@@

@@

## Plot effect of probability cut-offs

```{r , warning=FALSE}

tprs.plot <- tprs %>% ungroup() %>% select(probability,E6,E12) %>% pivot_longer(cols=c(E6,E12)) %>% mutate(name=gsub("^E",'',name))

size <- 2
p <- ggplot(tprs.plot)
p <- p + geom_line(aes(x=probability,y=value,col=name)) + ylab("Cost effect (€ / donation)") + xlab("Probability of deferral cut-off") + labs(color="Deferral length (mo)")
p <- p + geom_point(aes( y =tprs$E6[which(tprs$E6 == min(tprs$E6))] , x = tprs$probability[which(tprs$E6 == min(tprs$E6))] ))
p <- p + geom_label(aes( y =tprs$E6[which(tprs$E6 == min(tprs$E6))] , x = tprs$probability[which(tprs$E6 == min(tprs$E6))] , label = paste0("p>=",round(tprs$probability[which(tprs$E6 == min(tprs$E6))],2) , " ; ",  round(tprs$E6[which(tprs$E6 == min(tprs$E6))],2)," €")), nudge_y=0.2,size=size)
p <- p + geom_point(aes( y =tprs$E12[which(tprs$E12 == min(tprs$E12))] , x = tprs$probability[which(tprs$E12 == min(tprs$E12))] ))
p <- p + geom_label(aes( y =tprs$E12[which(tprs$E12 == min(tprs$E12))] , x = tprs$probability[which(tprs$E12 == min(tprs$E12))] , label = paste0("p>=",round(tprs$probability[which(tprs$E12 == min(tprs$E12))],2) , " ; ",  round(tprs$E12[which(tprs$E12 == min(tprs$E12))],2)," €")), nudge_y=0.2,size=size)
p <- p + theme(legend.position="bottom")
p
ggsave(filename="../results/cme_prob_cut_curve.pdf", p, width = 90,  height = 80,units="mm", dpi=600, scale=1.0,encoding = "ISOLatin9.enc")
```

Using a bit higher than the default cut-off  appears slightly better.

# Bootstrap cost confidence intervals

Let's bootstrap predictions to create error estimates for the predictions and subsequently on the cost effects.

```{r}
#Define functions
#boot code from https://github.com/FRCBS/changes_in_donor_health/blob/master/src/hypothesis_regressions.Rmd
boot_cost <- function(boot_data, boot_ind, prob.cut = 0.6){
  #sample
    boot_data <- boot_data[boot_ind,]
  #classify predictions
    pred.class <- factor( boot_data$Deferred >= prob.cut)
    levels(pred.class ) <- c("Accepted","Deferred")
  #get TPR and FPR
    conf <- caret::confusionMatrix(reference=boot_data$obs, data=pred.class, positive = "Deferred",mode="sens_spec")
    sen <- conf$byClass['Sensitivity'] 
    spe <- 1- conf$byClass['Specificity'] 
  #get cost
#    cat(sen,spe,"\n")
     c <- get_cost(TPR=sen, FPR=spe,d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl)
     c <- as.vector(t(c))
     names(c) <- c('q','a6','a12','E6','E12')
  return(c)
}

compute_CI <-function(fit_boot){
  CI_inf = rep(0, length(fit_boot$t0))
  CI_sup = rep(0, length(fit_boot$t0))
  for (i_regressor in 1:length(fit_boot$t0)){
    #https://stackoverflow.com/questions/6791290/r-boot-package-not-enough-memory-to-get-confidence-intervals
    #CI <- boot.ci(fit_boot, type = "bca", index=i_regressor)
    #Bca_inf[i_regressor] <- CI$bca[4] # N.B type = 'normal' has 3 columns, while the other types 5 columns
    #Bca_sup[i_regressor] <- CI$bca[5] #
    #there is just too much data for bca to be used
    CI <- boot.ci(fit_boot, type = "norm", index=i_regressor)
    CI_inf[i_regressor] <- CI$normal[2]
    CI_sup[i_regressor] <- CI$normal[3]
  }
  return(tibble(variable = names(fit_boot$t0),CI_inf,CI_sup))
}
```

@@

#nb_boot <- 1000
nb_boot <- nrow(df)
fit_boot <- boot(df,prob.cut=0.6, statistic= boot_cost, R = nb_boot)
p6.ci <- compute_CI(fit_boot)

fit_boot <- boot(df,prob.cut=0.8, statistic= boot_cost, R = nb_boot)
p8.ci <- compute_CI(fit_boot)

p6.ci
```

```{r}
p8.ci
```

## Plot surface with confidence intervals

```{r ,warning=FALSE}
#Get the a (y) and q (x) mappings for the optimised 6 and 12 month scenarios
x6 <- head(which(q.seq > tprs$q[which(tprs$E6 == min(tprs$E6))] ),1)
y6 <- head(which(adjustments > tprs$a6[which(tprs$E6 == min(tprs$E6))]),1)
x12 <- head(which(q.seq > tprs$q[which(tprs$E12 == min(tprs$E12))] ),1)
y12 <- head(which(adjustments > tprs$a12[which(tprs$E12 == min(tprs$E12))]),1)
#Plot them
colnames(E.long) <- c("adjustment", "q", "cost")
p <- ggplot(E.long, aes(x = q, y = adjustment, z = cost)) +
  geom_tile(aes(fill=cost)) +
  stat_contour(color = "white", size = 0.7, binwidth =  1) +
  scale_fill_gradientn(colours=c("blue", "red")) +
  labs(x="q% i.e. deferral rate", y=expression(a[tot]~i.e.~donation~interval~coefficient ))  + 
  theme_bw() +
  scale_x_continuous(breaks = c(1, 1001), labels = c("0", "3.2")) +
  scale_y_continuous(breaks = c(1, 1001), labels = c(as.character(min(adjustments)), as.character(max(adjustments))))
#Add numeric values of costs with optimised classifier probability cut-offs
p <- p + annotate("text" ,
                  x = x6 , 
                  y = y6 - 50,
                  label = paste0("6 mo: ", round(tprs$E6[which(tprs$E6 == min(tprs$E6))],2),' (',round(deframe(p6.ci[4,2]),2)," - ",round(deframe(p6.ci[4,3]),2),') €'  ),
                  colour = "white")

p <- p + annotate("text" ,
                  x = x12 , 
                  y = y12 - 50,
                  label = paste0("12 mo: ", round(tprs$E12[which(tprs$E12 == min(tprs$E12))],2),' (',round(deframe(p8.ci[5,2]),2)," - ",round(deframe(p8.ci[5,3]),2),') €'  ),
                  colour = "white")
#Add points estimates
p <- p + annotate("point", x = x6, y = y6, colour = "white", size = 3, alpha = 1) 
p <- p +  annotate("point", x = x12, y = y12, colour = "white", size = 3, alpha = 1)

#Get x and y values for error bars
xmin.p6  <- head(which(q.seq >  deframe(p6.ci[1,'CI_inf'])) ,1) 
xmax.p6  <- head(which(q.seq >  deframe(p6.ci[1,'CI_sup'])) ,1) 
ymin.p6  <- head(which(adjustments >  deframe(p6.ci[2,'CI_inf'])) ,1) 
ymax.p6  <- head(which(adjustments >  deframe(p6.ci[2,'CI_sup'])) ,1)

xmin.p12  <- head(which(q.seq >  deframe(p8.ci[1,'CI_inf'])) ,1) 
xmax.p12  <- head(which(q.seq >  deframe(p8.ci[1,'CI_sup'])) ,1) 
ymin.p12  <- head(which(adjustments >  deframe(p8.ci[3,'CI_inf'])) ,1) 
ymax.p12  <- head(which(adjustments >  deframe(p8.ci[3,'CI_sup'])) ,1) 
#Add error bars of q and a

p <- p + annotate("errorbar",x=x6,ymin = ymin.p6, ymax= ymax.p6,size=1   )
p <- p + annotate("errorbarh",y=y6,xmin = xmin.p6, xmax= xmax.p6,size=1   )

p <- p + annotate("errorbar",x=x12,ymin = ymin.p12, ymax= ymax.p12,size=1  )
p <- p + annotate("errorbarh",y=y12,xmin = xmin.p12, xmax= xmax.p12,size=1   )

p <- p + labs(fill="Cost effect (€ / donation)")  + theme(legend.position="bottom")
p
ggsave(filename="../results/cme_cost_surface_werr.pdf",  width = 180,  height = 120,units="mm", dpi=600, scale=1.0,encoding = "ISOLatin9.enc")
```
